---
title: "Base R 1: Vectors"
author: "Maria Ren"
date: "January 22, 2018"
header-includes:
  \DeclareMathSizes{12}{12}{12}{12}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Create the vectors:

 (a) (1, 2, 3, . . . , 19, 20)  
 
******
 
```{r}

1:20

```
******

 (b) (20, 19, . . . , 2, 1)
```{r}
c(20:1)
```

 (c) (1, 2, 3, . . . , 19, 20, 19, 18, . . . , 2, 1)
```{r}

c(1:20,19:1)
```

 (d) assign vector c(4, 6, 3) variable name tmp

 Use tmp for parts (e), (f) and (g)
 
******
```{r}
tmp <- c(4,6,3)

```
******
 (e) (4, 6, 3, 4, 6, 3, . . .  , 4, 6, 3) where there are 10 occurrences of 4.
```{r}
rep(tmp,times=10)
```

 (f) (4, 6, 3, 4, 6, 3, . . . , 4, 6, 3, 4) where there are 11 occurrences of 4, 10 occurrences of 6 and 10 occurrences of 3.

```{r}
rep(tmp,times=11)[1:31]
```
 (g) (4, 4, . . . , 4, 6, 6, . . .  , 6, 3, 3, . . . , 3) where there are 10 occurrences of 4, 20 occurrences of 6 and 30 occurrences of 3.
 
 ******
```{r}
rep(tmp, times=c(10,20,30))

```
******
### 2. Create a vector of the values of
$e^x cos(x)$ at x = 3, 3.1, 3.2, . . . ,6.

******
```{r}
# reset value of tmp

tmp <- seq(3,6, by = .01)

tmp1 <- exp(tmp)*cos(tmp)

head(tmp1, n = 30)

# To keep the listing shorter

```
******

### 3. Create the following vectors:

 (a) $(0.1^3 0.1^1, 0.1^6 0.2^4, ... ,0.1^{36} 0.2{34})$
```{r}
x <- seq(3,36,by=3)
y <- seq(1,34,by=3)
c((0.1^x)*(0.2^y))

```
 (b) $(2, \frac{2^2}{2}, \frac{2^3}{3}, ... ,\frac{2^{25}}{25})$
 
 ******
```{r}
(2^(1:25))/(1:25)
```
******

### 4. Calculate the following:

 (a) $\sum\limits_{i = 10}^{100}(i^3 + 4i^2)$
```{r}
i <- seq(10,100,by=1)
sum((i^3)+(4*(i^2)))
```
 (b) $\sum\limits_{i=1}^{25}(\frac{2^i}{i}+\frac{3^i}{i^2})$
 
 ******
```{r}
 
 tmp = 1:25
 
 sum((2^tmp)/tmp + 3^tmp/(tmp^2))
 
```



### 5. Use the function paste() to create the following character vectors of length 30:

 (a) ("label 1", "label 2", ....., "label 30").
Note that there is a single space between label and the number following.
```{r}
y <- paste(c("label"),1:30, sep=" ")
c(y)
```
 (b) ("fn1", "fn2", ..., "fn30").
```{r}
paste(c("fn"),1:30,sep="")

```
In this case, there is no space between fn and the number following.
******
```{r}
paste("fn", 1:30, sep="" )

```

******

### 6. Execute the following lines which create two vectors of random integers which are chosen with replacement from the integers 0, 1, . . . , 999. Both vectors have length 250.

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```
******

Suppose $x = (x_1, x_2, . . . ,x_n)$ denotes the vector xVec and
$y = (y_1, y_2, . .  ,y_n)$ denotes the vector yVec.


 (a) Create the vector
$(y_2 - x_1, . . . , y_n - x_{n-1})$.
```{r}
y <- yVec[-1]
n <- seq((1:length(y))-1)
y-xVec[n]
```
 (b) Create the vector
$(\frac{sin(y_1)}{cos(x_2)}, \frac{sin(y_2)}{x_3}, ...,\frac{y_{n-1}}{cos(x_n)})$.

```{r}
tmp1 <- sin(yVec[-length(yVec)])/cos(xVec[-1])

head(tmp1, n = 30)

```
******


 (c) Create the vector
$(x_1 + 2x_2 - x_3, x_2 + 2x_3 - x_4, ..., x_{x-2} + 2x_{n-1} - x_n)$
```{r}
first <- xVec[1:(length(xVec)-2)]
second <- xVec[2:(length(xVec)-1)]
third <- xVec[3:length(xVec)]
first+2*second-third 
```
 (d) Calculate
$\sum\limits_{i=1}^{n-1} \frac{e^{-x_{i+1}}}{x_i + 10}$

```{r}
a <- 2:(length(xVec))
b <- 1:(length(xVec)-1)
sum(exp(-xVec[a])/(xVec[b]+10))

```


### 7. This question uses the vectors xVec and yVec created in the previous question and the functions sort,
order, mean, sqrt, sum, and abs.

 (a) Pick out the values in yVec which are > 600.
```{r}
yVec[yVec>600]
```
 (b) What are the index positions in yVec of the values which are > 600?
```{r}
index <- 1:length(yVec)
index[yVec>600]
```
 (c) What are the values in xVec which correspond to the values in yVec which are > 600? (By correspond, we mean at the same index positions.)
```{r}
index <- 1:length(yVec)
num <- index[yVec>600]
xVec[num]

```

 (d) Create the vector $(|x_1 - \bar{x}|^{1/2}, |x_2 - \bar{x}|^{1/2}, ..., |x_n - \bar{x}|^{1/2})$
```{r}
x_bar <- mean(xVec)
(abs(xVec-x_bar))^(1/2)
```
 (e) How many values in yVec are within 200 of the maximum value of the terms in yVec?
```{r}
max <- max(yVec)
length(yVec[yVec>(max-200) & yVec<(max+200)])

```
 (f) How many numbers in xVec are divisible by 2? (Note that the modulo operator is denoted \%\%.)
```{r}
length(xVec[xVec%%2==0])
```
 (g) Sort the numbers in the vector xVec in the order of increasing values in yVec.
 
```{r}
 xVec[order(yVec)]
```
******

 (h) Pick out the elements in yVec at index positions 1, 4, 7, 10, 13, . . .
```{r}
index <- seq(1,250,by=3)
yVec[index]
```

## 8. By using the function cumprod or otherwise, calculate

$1 + \frac{2}{3} + (\frac{2}{3}\frac{4}{5}) + (\frac{2}{3}\frac{4}{5}\frac{6}{7} + ... + (\frac{2}{3}\frac{4}{5} ... \frac{38}{39}))$
```{r}
a <- seq(2,38,by=2)
b <- seq(3,39,by=2)
c <- seq(length(a))
1+sum(cumprod(a[c])/cumprod(b[c]))
```
