---
title: "Simple R Functions "
author: "Maria Ren"
date: "January 26, 2018"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}



c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$

```{r}
tmpFn3 <- function(x,n){
  
  z <- c(1:n)

  return(1+sum(x^z/z))
}


# simple example
tmpFn3(3,2)

```
\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

Try out your function.  \texttt{tmpFn}(c(1:5,6:1))

```{r}

tmpFn <- function(xVec){

a <- c(3:length(xVec))
y <- (xVec[a-2]+xVec[a-1]+xVec[a])/3

return(y)
}

# Simple example 
tmpFn(c(1:5,6:1))

```
### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.

```{r}

tmpFn <- function(xVec) {

  if (xVec < 0) {
     z <- (xVec ^2 + 2*xVec + 3)
     } else if (xVec < 2) {
     z <- (xVec + 3)
      } else if (xVec > 2) {
        z <- (xVec^2 + 4*xVec-7)
      }
    z
}


# Simple example

tmpFn(6)
tmpFn(-6)

```

plot the function tmpFn

```{r}
tmpFn <- function(xVec) {
    ifelse (xVec < 0, xVec ^2 + 2*xVec + 3,
        ifelse (xVec< 2, xVec + 3, 
        xVec^2 + 4*xVec-7))
}
    plot(tmpFn,xlim = c(-3,3))
```
\newpage

### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$

```{r}
func <- function(x) {
  x[x%%2==1] <- 2*x[x%%2==1]
  x
  
}



#Simple sample
a <- matrix(c(1,5,-2,1,2,-1,3,6,-3),nrow=3)
func(a)


```

### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$

```{r}
func <- function(n,k) {
  a <- matrix(0*(1:n),nrow=n,ncol=n)
  diag(a) <- k
  a[abs(row(a)-col(a))== 1] <-  1
  a
  
}


#Simple example

func(5,2)

```
\newpage
### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.

```{r}
quadrant <- function(alpha){
  a <- alpha%%360
   if (a< 90) {
     return ("quadrant 1")
  }else if (a<180) {
    return ("quadrant 2")
  }else if (a<270) {
     return ("quadrant 3")
  }else {
    return ("quadrant 4")
  }
}



#Simple Sample 
quadrant (77)
quadrant (166)
quadrant (224)
quadrant (280)
quadrant (390)




```
### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.

```{r}
weekday <- function(day,month,year){
    month <- month - 2
   if (month < 1) {
     month <- month + 12
     year <- year -1
   }
  y <- year%%100
  c <- (year - y)/100
  f <- (floor(2.6*month-0.2)+day+y+floor(y/4)+floor(c/4)-(2*c))%%7

c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")[f+1]
}


# Trials
weekday(8,2,2018)
```
\newpage
(b) Does your function work if the input parameters \texttt{day, month, and year} are vectors with the same length and valid entries?


```{r}
# Try with vectors
#weekday(c(8,9,2),c(3,3,5),(2018,1995,1854))

#Error: unexpected ',' in "weekday(c(8,9,2),c(3,3,5),(2018,"

#As seen above, My function would not work if the input parameters are vectors with the same length and valid entries, as shown above, it creates an error message. 

```


### 8. 

(a) Suppose $x_0 =1$ and $x_1 =2$ and

$$ x_{j}=x_{j - i}+\frac{2}{x_{j - 1}} $$
for $j=1,2,...$
Write a function \texttt testLoop which takes the single argument $n$ and returns the first $n - 1$ values of the sequence $x_j$ $j > 0$: that means the values of $x_0$, $x_1$, $x_2$, . . . , $x_{n} - 2$.

```{r}
testloop <- function(n) {
  x <- rep(0,length=n-1)
  i <- c(0:(n-2))
  x[1] <- 1
  x[2] <- 2
  for (i in 3:(n-1))
    x[i] <- x[i-1]+(2/x[i-1])
  x
}


          
#simple test 
testloop(5)
  

```
Now write a function \texttt testLoop2 which takes a single argument \texttt yVec which is a vector. The function should return

$$\sum_{j=1}^{n} e^{j}$$
where $n$ is the length of \texttt yVec.

```{r}
testloop2 <- function(yVec){
  n <- length(yVec)
  x <- 1:n
  sum(exp(x))
}


#simple test
testloop2(c(4,6,7))


```


### 9. 

Solution of the difference equation $$x_{n}=rx_{n - 1}(1 - x_{n - 1})$$ with starting value $x_1$.

(a) Write a function $quadmap( start, rho, niter )$ which returns the vector $(x_1,..., x_n)$ where $$x_{k}=rx_{k - 1}(1 - x_{k - 1})$$ and
niter denotes $n$, 
start denotes $x_1$, 
and rho denotes $r$.


Try out the function you have written:
for $r=2$ and $0 < x_1$ you should get $x_n \rightarrow 0.5$ as $n \rightarrow infinity$. 
try $tmp \leftarrow quadmap(start=0.95, rho=2.99, niter=500)$
Now switch back to the Commands window and type:
    $plot(tmp, type= "l")$
Also try the plot $plot(tmp[300:500], type="l")$

```{r}
quadmap <- function(start,rho,niter){
  i <- c(1:niter)
  x <- rep(0,length=niter)
  x[1] <- start
  for (i in (2:niter))
    x[i] <- rho*x[i-1]*(1-x[i-1])
  x
}



#simple test
quadmap(2,3,4)



# test the question asked for
quadmap(0.3,2,50)
tmp <- quadmap(start=0.95,rho=2.99,niter=500)
plot(tmp,type='l')
plot(tmp[300:500],type = 'l')


 
```


(b)Now write a function which determines the number of iterations needed to get $|x_n - x_{(n - 1)}| < 0.02$. So this function has only 2 arguments: start and rho. (For $start=0.95$ and $rho=2.99$, the answer is 84.)


```{r}
quadmapnew <- function(start,rho){
  a <- start
  b <- rho*a*(1-a)
  count <- 1
  while (abs(b-a) >= 0.02) {
    a <- b
    b <-  rho*a*(1-a)
    count <- count + 1
  }
   count 
    
    
}

#Simple test
quadmapnew(0.95,2.99) 

```



### 10. 

Given a vector $(x_1,..., x_n)$, the sample autocorrelation of $lag k$ is defined to be
$$r_{k}=\frac{\sum_{i=k+1}^{n}(x_{i} - \bar{x})(x_{i - k} - \bar{x})}{\sum_{i=1}^{n}(x_{i} - \bar{x})^2}$$
Thus
$$r_{1}=\frac{\sum_{i=2}^{n}(x_{i}-\bar{x})(x_{i-1} - \bar{x})}{\sum_{i=1}^{n}(x_{i} - \bar{x})^2} =\frac{(x_{2} - \bar{x})(x_{1} - \bar{x}) + ...(x_{n} - \bar{x})(x_{n - 1} - \bar{x})}{\sum_{i=1}^{n}(x_{i} - \bar{x})^2}$$
Write a function $tmpFn(xVec)$ which takes a single argument xVec which is a vector and returns a list of two values: $r_1$ and $r_2$.
In particular, find $r_1$ and $r_2$ for the vector $(2, 5, 8,..., 53, 56)$.

```{r}

  
tmpFn <- function(xVec){
   n <- length(xVec)
   m <- mean(xVec)
   a <- sum((xVec[2:n]-m)*
              (xVec[1:(n-1)]-m))
   b <- sum((xVec[1:n]-m)^2)
   r1 <- a/b


   e <- sum((xVec[3:n]-m)*
              (xVec[1:(n-2)]-m))
   f <- sum((xVec[1:n]-m)^2)
   r2 <- e/f

  list(r1,r2)
}


 #simple test
 v <- 1:19
 z <- 3*v-1
 tmpFn(z)

```

(b)(Harder.) Generalise the function so that it takes two arguments: the vector $xVec$ and an integer $k$ which lies between $1$ and $n -1$ where $n$ is the length of xVec.
The function should return a vector of the values $(r_0 = 1, r_1 , . . . , r_k )$.
If you used a loop to answer part (b), then you need to be aware that much, much better solutions are possible see exercises 4. (Hint: sapply.)



```{r}
tmpFn1 <- function(xVec,k){
  n <- length(xVec)
  m <- mean(xVec)
  tmpFn2 <- function(a){
  w <- sum((xVec[(a+1):n]-m)*(xVec[1:(n-a)]-m))
  z <- sum((xVec[1:n]-m)^2)
  w/z
  }
  r0 <- 1
  r <- sapply((1:k),tmpFn2)
  c(r0,r)
}


#simple test 

tmpFn1(c(1,2,3,4,6,7,8,9,3,9),3)

```















